---
title: "SolarAPI Data Collection"
author: "Henry Strecker"
date: "2025-02-03"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(httr)
library(jsonlite)
library(tidyverse)
library(readxl)
```

## Read in data sources

```{r}
# Read in data and panel model
fqhc <- read.csv('data/data_layers_jan20.csv')

panels_model <- readRDS('data/panel_count_model.rds')

api_key <- 'REDACTED'

```


```{r}
# Define a function similar to previous iterations that focuses on using the panel model to generate SolarAPI responses for each site
fetch_solarapi <- function(latitude, longitude, site_id, key){
  Sys.sleep(0.11) # Sleep for a fraction of a second so as to not to exceed the API request limit (600/min)
  
  # Call the API to get results for a given site
  solar_api_test <- GET(paste0('https://solar.googleapis.com/v1/buildingInsights:findClosest?location.latitude=', latitude, 
                               '&location.longitude=', longitude, '&requiredQuality=HIGH&key=', key))
  
  # Convert JSON to tabular format
  test_data <- fromJSON(rawToChar(solar_api_test$content))
  
  # Check if the requiredQuality=HIGH parameter was met, if not run requiredQuality=MEDIUM
  if (!is.null(test_data$error$message) && test_data$error$message == 'Requested entity was not found.'){
    solar_api_test <- GET(paste0('https://solar.googleapis.com/v1/buildingInsights:findClosest?location.latitude=', latitude, 
                                 '&location.longitude=', longitude, '&requiredQuality=MEDIUM&key=', key))
  
    # Convert JSON to tabular format
    test_data <- fromJSON(rawToChar(solar_api_test$content))
  }
  
  # Check if the requiredQuality=MEDIUM parameter was met, if not run requiredQuality=LOW
  if (!is.null(test_data$error$message) && test_data$error$message == 'Requested entity was not found.'){
    solar_api_test <- GET(paste0('https://solar.googleapis.com/v1/buildingInsights:findClosest?location.latitude=', latitude, 
                                 '&location.longitude=', longitude, '&requiredQuality=LOW&key=', key))
  
    # Convert JSON to tabular format
    test_data <- fromJSON(rawToChar(solar_api_test$content))
  }
  
  # Deal with edge case where API doesn't pass due to an API restriction
  while (!is.null(test_data$error$message) && grepl('IP address restriction', test_data$error$message)){
    solar_api_test <- GET(paste0('https://solar.googleapis.com/v1/buildingInsights:findClosest?location.latitude=', latitude, 
                                 '&location.longitude=', longitude, '&requiredQuality=LOW&key=', key))
  
    # Convert JSON to tabular format
    test_data <- fromJSON(rawToChar(solar_api_test$content))
  }
  
  # Write SolarAPI output to save all information
  write_rds(test_data, paste0('data/solarapi_responses/', 'site_', site_id, '_solarapi.rds'))
  
  # Handle case where there are insufficient/missing SolarAPI results - no data returned here
  if (is.null(test_data$solarPotential) & length(test_data$solarPotential$maxArrayPanelsCount) == 0){
    return(rep(NA, 8)) # In case SolarAPI doesn't produce results for a site
  } else if (is.null(test_data$solarPotential$maxArrayPanelsCount) || test_data$solarPotential$maxArrayPanelsCount < 10){
    return(rep(NA, 8)) # In case SolarAPI can't place a minimum of 10 panels on a roof
  }
  
  # Maximum number of panels from api call
  max_panels <- as.numeric(test_data$solarPotential$maxArrayPanelsCount)
  
  # Total roof area in square feet from api call converted from square meters
  roof_area_sqft <- test_data$solarPotential$wholeRoofStats$areaMeters2 * 10.7639
  roof_area_sqm <- test_data$solarPotential$wholeRoofStats$areaMeters2
  
  # Store the master set of panel configurations and solar output for this site
  panels_generation <- test_data$solarPotential$solarPanelConfigs
  
  # Define roof size indicator variables for the panel modeling prediction
  if (roof_area_sqft <= 7000){
    small_area <- 1
    medium_area <- 0
  } else if (roof_area_sqft <= 15000){
    small_area <- 0
    medium_area <- 1
  } else {
    small_area <- 0
    medium_area <- 0
  }
  
  # Estimate true number of panels based on max_panels and roof area using the linear regression model
  panel_prediction <- unname(predict(panels_model, list(max_panel_count = max_panels, small_site = small_area, medium_site = medium_area)))
  
  # Extracts the configuration that matches the number of panels modeled above
  annual_solar_generation <- panels_generation[panels_generation$panelsCount == panel_prediction, ]
  panel_count <- panel_prediction

  # If the number of modeled panels doesn't exactly match with the API call, this will find the next closest match
  if (dim(annual_solar_generation)[1] == 0){
    panel_count <- panels_generation$panelsCount[which.min(abs(panels_generation$panelsCount - panel_prediction))]
    annual_solar_generation <- panels_generation[panels_generation$panelsCount == panel_count, ]
  }
  
  nameplate_capacity <- panel_count * 400 / 1000 # 400W panels, scaled to kW from W
  
  # Generate true solar production potential, based on 400W panels and solar generation similar to Aurora screenings (scale at 88.5%)
  adjusted_solar_annual_kwh <- annual_solar_generation$yearlyEnergyDcKwh * (0.885)
  
  # Relevant number of roof segments - meant for solving issue with residential buildings that have roof segments that can't house panels
  roof_segment_count <- as.numeric(length(pull(tail(test_data$solarPotential$solarPanelConfigs$roofSegmentSummaries, 1)[[1]]['panelsCount'])))
  
  # Calculate roof height ine meters above sea level, taking the average across all roof segments
  building_height_sealevel <- mean(test_data$solarPotential$roofSegmentStats$planeHeightAtCenterMeters)
  
  # Extract median irradiance for each roof segment
  median_irradiance <- c()
  for (i in 1 : roof_segment_count){
    n_breaks <- length(test_data$solarPotential$roofSegmentStats$stats$sunshineQuantiles[[i]]) # Number of breaks for percentile calculations
    median_irradiance <- c(median_irradiance, test_data$solarPotential$roofSegmentStats$stats$sunshineQuantiles[[i]][n_breaks %/% 2 + 1]) # Ensures median is extracted
  }
  
  # Create roof area data for how much of the roof faces each direction
  roof_stats <- test_data$solarPotential$roofSegmentStats
  roof_stats_df <- data.frame(pitch = head(roof_stats$pitchDegrees, roof_segment_count),
                              azimuth = head(roof_stats$azimuthDegrees, roof_segment_count),
                              roof_area_sqm = head(roof_stats$stats$areaMeters2, roof_segment_count),
                              roof_segment_panel_count = pull(tail(test_data$solarPotential$solarPanelConfigs$roofSegmentSummaries, 1)[[1]]['panelsCount']),
                              roof_segment_irradiance = median_irradiance)
  
  solar_panel_size_sqm <- test_data$solarPotential$panelHeightMeters * test_data$solarPotential$panelWidthMeters
  num_roof_segments <- as.numeric(dim(roof_stats_df)[1])
  
  # Initialize variables
  usable_roof_area_sqm <- 0
  irradiance_times_area <- 0
  
  # Calculate total roof summary statistics
  for (i in 1:dim(roof_stats_df)[1]){
    if (roof_stats_df$pitch[i] <= 10){ # Included because flat roofs are given azimuth of 0
      usable_roof_area_sqm <- usable_roof_area_sqm + roof_stats_df$roof_segment_panel_count[i] * solar_panel_size_sqm
      irradiance_times_area <- irradiance_times_area + (roof_stats_df$roof_segment_panel_count[i] * solar_panel_size_sqm * roof_stats_df$roof_segment_irradiance[i])
    } else if (roof_stats_df$pitch[i] <= 45 & roof_stats_df$azimuth[i] < 315 & roof_stats_df$azimuth[i] > 45){
      usable_roof_area_sqm <- usable_roof_area_sqm + roof_stats_df$roof_segment_panel_count[i] * solar_panel_size_sqm
      irradiance_times_area <- irradiance_times_area + (roof_stats_df$roof_segment_panel_count[i] * solar_panel_size_sqm * roof_stats_df$roof_segment_irradiance[i])
    }
  }
  
  # Electricity output = solar irradiance * usable roof area * panel efficiency * performance ratio
  irradiance_method_solar_potential_kwh <- irradiance_times_area * 0.204 * 0.85
  
  ######## OLD VERSION OF GENERATING RESULTS
  # Initialize variables
  # flat_roof_area <- 0
  # south_roof_area <- 0
  # east_west_roof_area <- 0
  # north_roof_area <- 0
  # usable_flat_roof_area <- 0
  # usable_south_roof_area <- 0
  # usable_east_west_roof_area <- 0
  # usable_north_roof_area <- 0
  # 
  # 
  # Loop through the roof segments and calculate area totals
  # Flat roofing is defined as having pitch less than 10 degrees
  # Roofing with pitch greater than 45 degrees is disregarded
  # Azimuth (orientation) is defined with 0 as North, 180 as South
  # Directions are defined as 90 degree windows centered around the cardinal directions
  # For example, south is considered between 135 and 225 degrees
  # Code was perhaps too granular, we'll lump all flat, North, East, and West facing roofing together
  # for (i in 1:num_roof_segments){
  #   if (roof_stats_df$pitch[i] >= 45 ){
  #     next
  #   }
  #   if (roof_stats_df$pitch[i] <= 10){
  #     flat_roof_area <- flat_roof_area + roof_stats_df$roof_area[i]
  #     usable_flat_roof_area <- usable_flat_roof_area + roof_stats_df$roof_segment_panel_count[i] * solar_panel_size_sqft
  #   } else if (roof_stats_df$azimuth[i] > 135 && roof_stats_df$azimuth[i] <= 225){
  #     south_roof_area <- south_roof_area + roof_stats_df$roof_area[i]
  #     usable_south_roof_area <- usable_south_roof_area + roof_stats_df$roof_segment_panel_count[i] * solar_panel_size_sqft
  #   } else if (roof_stats_df$azimuth[i] > 45 && roof_stats_df$azimuth[i] <= 135 |
  #              roof_stats_df$azimuth[i] > 225 && roof_stats_df$azimuth[i] <= 315){
  #     east_west_roof_area <- east_west_roof_area + roof_stats_df$roof_area[i]
  #     usable_east_west_roof_area <- usable_east_west_roof_area + roof_stats_df$roof_segment_panel_count[i] * solar_panel_size_sqft
  #   } else if (roof_stats_df$azimuth[i] > 315 | roof_stats_df$azimuth[i] <= 45){
  #     north_roof_area <- north_roof_area + roof_stats_df$roof_area[i]
  #     usable_north_roof_area <- usable_north_roof_area + roof_stats_df$roof_segment_panel_count[i] * solar_panel_size_sqft
  #   }
  # }
  ########

  # Relay the calculated values as output
  return(c(roof_area_sqm, usable_roof_area_sqm, panel_count, nameplate_capacity, 
           adjusted_solar_annual_kwh, num_roof_segments, building_height_sealevel))
}

```


# Validation test code 2/18/2025

This code won't work with the above function anymore because it's been altered to write the API response using site ID, which some of these are missing. If you comment out that line of code, this below will work fine
```{r}

# validation_data <- read.csv('data/solar_api_validation_data.csv') %>%
#   filter(data_source == 'screening') %>%
#   select(-c(15:19)) %>% 
#   rename(site_lat = lat,
#          site_lon = lon)
# 
# validation_data <- validation_data %>%
#   rowwise() %>% # Process row-by-row
#   mutate(solar_api_results = list(fetch_solarapi(latitude = site_lat, longitude = site_lon, 
#                                                  key = api_key))) %>% # Make API call and record results
#   mutate(roof_area_sqm = solar_api_results[[1]],
#          usable_roof_area_sqm = solar_api_results[[2]],
#          aurora_panel_count = solar_api_results[[3]],
#          aurora_nameplate_capacity = solar_api_results[[4]],
#          aurora_annual_solar_generation_kwh = solar_api_results[[5]],
#          num_roof_segments = solar_api_results[[6]],
#          building_height_sealevel = solar_api_results[[7]]) %>% 
#   ungroup() %>%
#   select(-solar_api_results) # Remove column that originally held the data
  
```

Test results that would accompany the validation code above
```{r}
# test_results <- validation_data %>%
#   filter(!is.na(validation_data$roof_area_sqm)) %>%
#   select(site_address, site_lat, site_lon, roof_area, roof_area_sqm, usable_roof_area_sqm, num_roof_segments, panel_count, nameplate_capacity, annual_solar_generation, 
#          aurora_panel_count, aurora_annual_solar_generation_kwh,irradiance_method_solar_potential_kwh) %>%
#   mutate(panel_model_comparison = aurora_annual_solar_generation_kwh/annual_solar_generation,
#          irradiance_model_comparison = irradiance_method_solar_potential_kwh/annual_solar_generation)
# 
# print(c(mean(test_results$panel_model_comparison), mean(test_results$irradiance_model_comparison)))
# Panel modeling overestimated results on average by about  12 percent, the usable roof area method overestimated by 130 percent

```


# Packard Plan Test
```{r}
# Read in data for the Packard Plan Sites
first_batch <- read.csv('data/solarapi_batch_nov26/Sites_2024_1313894487973521005.csv')
second_batch <- read.csv('data/solarapi_batch_nov26/Centers_2023_8353317920455565290.csv')

first_batch_results <- first_batch %>%
  rename(site_lat = y_sites_rooftop_CEC, site_lon = x_sites_rooftop_CEC) %>%
  rowwise() %>% # Process row-by-row
  mutate(solar_api_results = list(fetch_solarapi(latitude = site_lat, longitude = site_lon, 
                                                 key = api_key))) %>% # Make API call and record results
  mutate(roof_area_sqm = solar_api_results[[1]],
         usable_roof_area_sqm = solar_api_results[[2]],
         aurora_panel_count = solar_api_results[[3]],
         aurora_nameplate_capacity = solar_api_results[[4]],
         aurora_annual_solar_generation_kwh = solar_api_results[[5]],
         num_roof_segments = solar_api_results[[6]],
         building_height_sealevel = solar_api_results[[7]]) %>% 
  ungroup() %>%
  select(-solar_api_results) # Remove column that originally held the data

second_batch_results <- second_batch %>%
  rename(site_lat = y_centers_rooftop_CEC, site_lon = x_centers_rooftop_CEC) %>%
  rowwise() %>% # Process row-by-row
  mutate(solar_api_results = list(fetch_solarapi(latitude = site_lat, longitude = site_lon, 
                                                 key = api_key))) %>% # Make API call and record results
  mutate(roof_area_sqm = solar_api_results[[1]],
         usable_roof_area_sqm = solar_api_results[[2]],
         aurora_panel_count = solar_api_results[[3]],
         aurora_nameplate_capacity = solar_api_results[[4]],
         aurora_annual_solar_generation_kwh = solar_api_results[[5]],
         num_roof_segments = solar_api_results[[6]],
         building_height_sealevel = solar_api_results[[7]]) %>% 
  ungroup() %>%
  select(-solar_api_results) # Remove column that originally held the data

write.csv(first_batch_results, 'data/solarapi_batch_feb23/Sites_2024_1313894487973521005_solarapi.csv')
write.csv(second_batch_results, 'data/solarapi_batch_feb23/Centers_2023_8353317920455565290_solarapi.csv')

```


# Batch API Calls

If you'd want to generate results in smaller batches, simply run through fewer iterations of the for loop at a time

```{r}
# for (i in c(1)){ # Will help to split this process into smaller chunks
#   
#   iteration_data <- fqhc[((i-1)*10 + 1):(i*10), ] # Standard index case, pulls 10 records at a time
#   
#   # Make API Call
#   iteration_results <- iteration_data %>%
#     rowwise() %>% # Process row-by-row
#     mutate(solar_api_results = list(fetch_solarapi(latitude = site_lat, longitude = site_lon, 
#                                                    site_id = ObjectId, key = api_key))) %>% # Make API call and record results
#     mutate(roof_area_sqm = solar_api_results[[1]],
#            usable_roof_area_sqm = solar_api_results[[2]],
#            aurora_panel_count = solar_api_results[[3]],
#            aurora_nameplate_capacity = solar_api_results[[4]],
#            aurora_annual_solar_generation_kwh = solar_api_results[[5]],
#            num_roof_segments = solar_api_results[[6]],
#            building_height_sealevel = solar_api_results[[7]]) %>% 
#     ungroup() %>%
#     select(-solar_api_results) # Remove column that originally held the data
#   
#   # Write results to a csv
#   write.csv(iteration_results, file.path('data/solarapi_results', paste0('fqhc_solarapi_', i, '.csv')), row.names = FALSE)
# }


```


# Generate SolarAPI Results for All Sites

```{r}
# This code chunk will run SolarAPI for all sites in the database, and save their API responses

for (i in 1 : as.integer(dim(fqhc)[1] / 100) + 1){ # Will help to split this process into smaller chunks
  if (i == as.integer(dim(fqhc)[1] / 100) + 1){
    iteration_data <- fqhc[((i-1)*100 + 1):dim(fqhc)[1], ] # Special case for indexing at the end
  } else {
    iteration_data <- fqhc[((i-1)*100 + 1):(i*100), ] # Standard case for indexing, pulls 400 records at a time
  }

  # Make API Call
  iteration_results <- iteration_data %>%
    rowwise() %>% # Process row-by-row
    mutate(solar_api_results = list(fetch_solarapi(latitude = site_lat, longitude = site_lon,
                                                   site_id = ObjectId, key = api_key))) %>% # Make API call and record results
    mutate(roof_area_sqm = solar_api_results[[1]],
           usable_roof_area_sqm = solar_api_results[[2]],
           aurora_panel_count = solar_api_results[[3]],
           aurora_nameplate_capacity = solar_api_results[[4]],
           aurora_annual_solar_generation_kwh = solar_api_results[[5]],
           num_roof_segments = solar_api_results[[6]],
           building_height_sealevel = solar_api_results[[7]]) %>%
    ungroup() %>%
    select(-solar_api_results) # Remove column that originally held the data

  # Write results to a csv
  write.csv(iteration_results, file.path('data/solarapi_results', paste0('fqhc_solarapi_', i, '.csv')), row.names = FALSE)
}


```


Run this code to append all new data to the overall FQHC data file
Does not require the above code if the data is already present

```{r}
# Initialize an empty data frame
solarapi_results_all <- data.frame()

# Populate with solarapi results
for (i in list.files(file.path('data/solarapi_results'))){
  solarapi_results_all <- rbind(solarapi_results_all, read.csv(file.path('data/solarapi_results', i)))
}

solarapi_results_all <- solarapi_results_all %>%
  select(ObjectId, roof_area_sqm, usable_roof_area_sqm, aurora_panel_count,
         aurora_nameplate_capacity, aurora_annual_solar_generation_kwh, 
         num_roof_segments, building_height_sealevel) %>% # Only care about these columns
  rename(solarapi_panel_count = aurora_panel_count,
         solarapi_nameplate_capacity = aurora_nameplate_capacity,
         solarapi_annual_solar_generation_kwh = aurora_annual_solar_generation_kwh)

write.csv(solarapi_results_all, 'data/solarapi_results_total.csv', row.names = FALSE)

fqhc_output <- read.csv('data/data_layers_jan20.csv')

# Join results with original data
fqhc_output <- fqhc_output %>%
  left_join(solarapi_results_all, by = 'ObjectId')

write.csv(fqhc_output, 'data/data_layers_mar1.csv', row.names = FALSE) # Write final output

```

